DP problems of interest

### 2012-09-28 docs/roadmap/problems.tex
Added matrix mul, init plan.

% ----------------------------------------------
\subsection{Convex polygon triangulation}
Problem: triangulating a polygon with least cost overall for added edges

Alphabets:\ul
\item Input: $\Sigma(S)= $ a matrix of costs for every possible edge in the polygon. The 
size of the input is $\frac{n^2}{2}$, where $n$ is the number of vertices in the polygon. 
\item Cost matrix: $\Sigma(M)=\mathbb{N}$. The sum of cost of edges is not "bounded".
\item Backtrack matrix: $\Sigma(B)=\{n\}$
\ule

Recurrence:
\[M_{(i,j)}= \max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} + c(i,k) | k = B_{(i,j)} \]

It is interesting to note that even in the sequential world, this problem is best solved 
by filling the diagonals, ie. computing sub-solutions for all polygons on size $k$ before
those of size $k+1$.

% ----------------------------------------------

@@

Recurrence:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	M_{(i+1,j-1)}+\omega(i,j) & SW\\
	M_{(i+1,j)} & S\\
	M_{(i,j-1)} & W\\
	\max_{i<k<j}M_{(i,k)}+M_{(k+1,j)} & k
\end{array}\right\} = B_{(i,j)} \]
With $\omega(i,j)=1$ if $i,j$ are complementary. 0 otherwise.

% ------------------------------------------------------------------------------------------------
\newpage
\subsection{Zuker folding}
XXX: we let this aside for the moment as the recurrence has too many unknown parameters
Problem: folding a RNA string $S$ over itself.

@@

Recurrence:
\[\begin{array}{rcl}
W_{(i,j)}&=&\min\left\{\begin{array}{l}
	W_{(i+1,j)}+b \\
	W_{(i,j-1)}+b \\
	V_{(i,j)}+\delta(S_i,S_j) \\
	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)}
\end{array}\right.\\\\

V_{(i,j)}&=&\min\left\{\begin{array}{ll}
	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
	eh(i,j)+b \qquad\qquad\, \text{otherwise} \\
	V_{(i+1,j-1)}+es(i,j) \\
	VBI(i,j) \\
	\min_{i<k<j-1}\{W_{(i+1,k)}+W_{(k+1,j-1)}\} +c
\end{array}\right.\\\\

VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c \\\\

F_{(j)}&=&\min\left\{\begin{array}{ll}
F_{(j-i)} \\
\min_{1\le i< j} (V_{(i,j)} + F_{(j-1)})
\end{array}\right. \text{(Free Energy)}\\\\

\end{array}\]

\subsection{Comments}

VBI stores scores for internal loops and bulges.
Notice in VBI, two moving indices $i', j'$. In general this gives a quadratic complexity for that particular recurrence, making the overall complexity of the algorithm $O(n^4)$. In practice (as read in Arpith's work), we have that $i' - i + j'- j < 30$, so the complexity is reduced to $O(n^3)$, albeit with a large constant.

%XXX: missing: eh(i,j), ebi(i,j,i',j'), es(i,j), ...

>>>>

% ----------------------------------------------
\newpage
\subsection{Matrix chain multiplication}
Problem: find an optimal parenthesizing of the multiplication of $n$ matrices $A_i$ of dimension $r_i \times c_i$.

Alphabets:\ul
\item Input: matrix size is defined as pairs of integers $(r_i,c_i)$.
\item Cost matrix: the cost is an integer (might blow up, use a float or a double?).
\item Backtrack matrix: $\Sigma(B)=\{stop\} \cup \{0..n\}$ with $n$ the input length.
\ule

Initialization (line=length of sequence, column=start):\ul
\item Cost matrix: $M_{(0,j)}=0$.
\item Backtrack matrix: $B_{(0,j)}=stop$.
\ule

Recurrence for the cost matrix:
\[M_{(i,j)}=\min_{i\le k<j}\left\{\begin{array}{l|l}
	M_{(i,k)}+M_{(k,j)}+r_i \cdot r_k \cdot col_j & k
\end{array}\right\}=B_{(i,j)} \]

@@

Recurrence:
\[M_{(i,j)}=\max\left\{\begin{array}{l|l}
	M_{(i+1,j-1)}+\omega(i,j) & SW\\
	M_{(i+1,j)} & S\\
	M_{(i,j-1)} & W\\
	\max_{i<k<j}M_{(i,k)}+W_{(k+1,j)} & k
\end{array}\right\} = B_{(i,j)} \]
With $\omega(i,j)=1$ if $i,j$ are complementary. 0 otherwise.

% ------------------------------------------------------------------------------------------------
%\newpage
\subsection{Zuker folding}
 XXX: we let this aside for the moment as the recurrence has too many unknown parameters
Problem: folding a RNA string $S$ over itself.

@@

Recurrence:
\[\begin{array}{rcl}
W_{(i,j)}&=&\min\left\{\begin{array}{l}
	W_{(i+1,j)}+b \\
	W_{(i,j-1)}+b \\
	V_{(i,j)}+\delta(S_i,S_j) \\
	\min_{i<k<j}W_{(i,k)}+W_{(k+1,j)}
\end{array}\right.\\\\
V_{(i,j)}&=&\min\left\{\begin{array}{ll}
	\infty \qquad\qquad\qquad\qquad {\rm if}(S_i,S_j) \text{ is not a base pair}\\
	eh(i,j)+b \qquad\qquad\, \text{otherwise} \\
	V_{(i+1,j-1)}+es(i,j) \\
	VBI(i,j) \\
	\min_{i<k<j-1}\{W_{(i+,k)}+W_{(k+1,j-1)}\} +c
\end{array}\right.\\\\
 F_{(j)}&=&\min\left\{\begin{array}{ll}
F_{(j-i)} \\
\min_{1\le i< j} (V_{(i,j)} + F_{(j-1)})
\end{array}\right. \text{(Free Energy)}\\\\
VBI(i,j)&=&\min_{i<i'<j'<j}\{V_{(i',j')}+ebi(i,j,i',j')\} +c
\end{array}\]
%
%XXX: missing: eh(i,j), ebi(i,j,i',j'), es(i,j), ...
